Mitigating object prior-bias from sparse-projection tomographic reconstructions

### 2022-02-16 paper/response_2022.tex

\section{Comments from AE}

\textcolor{blue}{\textit{While the reviewers have acknowledged
    improvements compared to the earlier manuscript, they have raised
    some important concerns about the manuscript in its current
    form. While addressing each of the reviewer specific comments,
    please pay specific attention to the following:}}

  \begin{itemize}
  \item \textcolor{blue}{\textit{Characterization of deep learning
        methods and its relation to the presented work}}

    We have now included a discussion of deep-learning (DL) methods in
    `Related Work', Section 2, of our paper, and have cited a few
    notable references pertaining to training neural networks to
    directly predict full-view sinogram and reconstruct the test
    object from its limited-view measurements. We also discuss how
    our work is complementary to learning-based teachniques and can
    fit alongside a deep learning framework, as seen in Figure.2 of the paper. Hence, the ability of DL to
    uncover new features from one or many example training images is not being disputed. Here the focus is on a new discrete algorithm that builds structural maps that can and does point to new features.

@@

\item \emph{ \ldots all steps \ldots} We also provide the details
  required.  We note that the TV reconstruction \emph{is} our pilot
  reconstruction and hence the computation times are equal.

  Having specified the exact compute times in our machine, we also
  note that the type of machine and the degree of parallelism
  available in the current times make these timing measurements,
  especially for iterative solutions, subjective.

\end{enumerate}

\begin{figure}[p]
    \begin{subfigure}[b]{0.19\linewidth}
        \includegraphics[width=\textwidth, height=0.2\textheight]{../images/thesis/okra/templateCropped_1.png}
\captionsetup{labelformat=empty}       
% \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{0.19\linewidth}
        \includegraphics[width=\textwidth, height=0.2\textheight]{../images/thesis/okra/templateCropped_2.png}
\captionsetup{labelformat=empty}
%        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.19\linewidth}
        \includegraphics[width=\textwidth, height=0.2\textheight]{../images/thesis/okra/templateCropped_3.png}
\captionsetup{labelformat=empty}
%        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.19\linewidth}
        \includegraphics[width=\textwidth, height=0.2\textheight]{../images/thesis/okra/templateCropped_4.png}
\captionsetup{labelformat=empty}
%        \caption{}
     \end{subfigure}
    \begin{subfigure}[b]{0.186\linewidth}
        \fcolorbox{yellow}{yellow}{\includegraphics[width=\textwidth, height=0.2\textheight]{../images/thesis/okra/testCropped.png}}
\captionsetup{labelformat=empty}
%        \caption{}
     \end{subfigure}
     \caption[Okra 3D dataset]{One slice each from the templates and
       one from the test volume (extreme right). In the regions marked
       in red and green, while all slices have deformities, the test
       has none.}
\label{fig:templates_test_okra}
%\end{figure}
%\begin{figure}[p]
\centering
\subcaptionbox{Test}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/testCropped.png}}\hfill
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/fdk_cropped.png}}\hfill
%\subcaptionbox{CS,no prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/cs_cropped.png}}\hfill
\subcaptionbox{Only object-\\prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/pca_cropped.png}}\hfill
\subcaptionbox{Spatially\\ varying object-\\ prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/prior_weighted_cropped.png}}
\caption[3D reconstruction of okra dataset]{Reconstruction from $10\%$
  views: (b) strong streak artefacts, %(c) blurred,
  (c) no new
  information detected (prior dominates -- the deformity from the prior
  shows up as a false positive) and (d) new information detected (no deformities
  corresponding to red and green regions) while simultaneously
  reducing streak artefacts.}
\label{fig:okra_3D_results}
%\end{figure}
%\begin{figure}[p]
\centering
\subcaptionbox{Test}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/zoomed/input_cropped.png}}\hfill
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/zoomed/FDK_cropped.png}}\hfill
%\subcaptionbox{Sparsity prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/zoomed/pca_cropped.png}}\hfill
\subcaptionbox{Only object-\\prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/zoomed/pca_cropped.png}}\hfill
\subcaptionbox{Spatially\\ varying object-\\ prior}{\includegraphics[width=0.19\columnwidth, height=0.2\textheight]{../images/thesis/okra/zoomed/weighted_prior_cropped.png}}
\caption[Zoomed results of Okra dataset]{Zoomed RoI of Fig.~\ref{fig:okra_3D_results} (b) strong streak artefacts, %(c) blurred,
  (c) no new  information detected (prior dominates -- the deformity from the prior
  shows up as a false positive) and (d) new information detected (no deformities).}
\label{fig:okra_zoomed_3D_results}
\end{figure}

@@

\vspace{0.5cm}\textcolor{blue}{\textbf{[R2-3]:} \textit{In several parts of the
    paper, the authors claim that existing object-prior based methods
    have problems, e.g.: "A common problem with these techniques is
    the strong influence of object-priors in the reconstruction of new
    regions in the test". At the moment, this hard claim is not
    supported in the paper by references or (even better) comparison
    results. It would be good to show, on one (or more) of the
    datasets used, that this is indeed the case, and that the proposed
    methods solves this problem.}}

\textbf{Response:} We have now included a reference (Pg.2, Line 3
under Sec.1-A) to a PhD thesis. This thesis discusses the problem of object-priors biasing the reconstruction of new test regions.
A sample result (Figures 1-3) from the thesis is shown here. It
illustrates how a direct application of object-prior can adversely
affect the reconstruction of new regions in the test.

@@

\vspace{0.5cm}\textcolor{blue}{\textbf{[R2-4]:} \textit{The discussion of deep learning methods in Section 7B is not very strong in my opinion. There are many existing techniques that have shown quite convincing results for reconstructing CT images with a low number of projections, even without taking into account any static parts across time. These methods rely on learning image-based correlations instead of the "given intensities of a voxel at various time instants in a longitudinal setting and partial measurements of the voxel at the current time, what will be the intensity of the same voxel at the current time instant?" given by the authors. These image-based methods often don't need 'hundreds of labeled data', sometimes being able to accurately learn from just one time step. In the paper, these methods should be discussed, and proper references should be included.}}\\

\textbf{Response}: We acknowledge that there has been significant work pertaining to training neural networks to directly predict full-view sinogram and reconstruction of the test object from its limited-view measurements. We have now included a discussion of deep-learning methods in `Related Work', Section 2, of our paper, and have now cited a few notable references. We also discuss how
    our work is complementary to learning-based teachniques and can
    fit alongside a deep learning framework, as seen in Figure.2 of the paper. Hence, the ability of DL to
    uncover new features from one or many example training images is not being disputed. Here the focus is on a new discrete algorithm that builds structural maps that can and does point to new features.

### 2022-02-16 paper/post_tci.tex
reorgnaized literature

In this paper, we present a technique that makes use of multiple
object priors, corresponding to high quality reconstructions of
similar objects across time, typically in a longitudinal study.  We
use these priors to improve the reconstruction of the test from
largely undersampled measurements. In particular, we define regions of
change between the test and the eigenspace spanned by previous high
quality object-priors in a purely data driven manner. Additionally,
our technique differentiates between genuine structural changes and
``changes'' that appear due to undersampling artefacts.  Besides this,
since we use a statistical model with multiple object-priors, we avoid
the problem of selecting an appropriate single object-prior
unlike~\cite{PICCS,Pourmorteza2015,Lee2012,pirple,Marjolein2016} and
also make use of the additional information available in multiple
priors. Our technique also has the advantage of not requiring any
expert-domain knowledge about the object being scanned.

 In the recent past, deep learning based techniques have also been widely used for limited view CT reconstruction without the use of any longitudinal object-prior. For example, in~\cite{anirudh}, a neural network is trained to predict the complete sinogram from an incomplete one obtained from few-view imaging. This complete sinogram is then fed to analytical reconstruction algorithms to obtain the final reconstruction. Hence, one may in principle use this full sinogram as an alternate initial reconstruction of the test within our framework. \cite{bo2021} presented a cascade of neural networks that predict full-view reconstruction from limited view FBP reconstruction while simlutaneously ensuring consistency between the predicted sinugram and the acquired sinogram. \cite{special} trains an adverserial network for artifact removal and edge preservation in limited view reconstruction. While all of the above methods predict full-view reconstruction from the measurements of test object alone, we believe it is beneficial to use information from longitudinal data, especially if naturally available, like in the radio-frequency ablation datasets that we use. In this context, we present an analytical method that uses the available longitudinal data to better inform the reconstruction of the test. We first construct a `weights map' that gives clear insights about regions of change within the object over time.

@@

In this section, we first illustrate the advantage of using multiple
types of eigenspaces for the computation of weights map, and then
present reconstruction results on all datasets.  We show both
qualitative and quantitative results.  Before we do this, it is import
to first consider our method of reporting our results.

>>>>

In the recent past, deep learning based
techniques~\cite{anirudh},\cite{kim2019extreme},\cite{bo2021},\cite{special}
have been used for limited view CT reconstruction. For example,
in~\cite{anirudh}, a neural network is trained to predict the complete
sinogram (measurements) from an incomplete one.  This complete
sinogram is then fed to analytical reconstruction algorithms to obtain
the final reconstruction.  ~\cite{bo2021} presented a cascade of
neural networks that predict full-view reconstruction from limited
view FBP reconstruction while simultaneously ensuring consistency
between the predicted sinogram and the acquired
sinogram. \cite{special} trains an adversarial network for artifact
removal and edge preservation in limited view reconstruction.
However, all of the above existing methods predict full-view
reconstruction from the measurements of the test object alone. We
believe it is beneficial to use information from longitudinal data,
especially if naturally available, like in the radio-frequency
ablation datasets that we use.

That said, the ability of these data-driven techniques to uncover new
features is not being disputed in this work. In fact, our work can be
used in tandem with these methods. For example, one of the state of
the art works is \cite{kim2019extreme}, where an adversarial network
is used to predict a reconstruction from extremely sparse measurements
(similar to the ones we employ in this work, albeit for chest
X-rays). This reconstruction is used as an initialization in a
traditional analytical iterative reconstruction algorithm to compute
the final reconstruction.  Given this pipeline, and our emphasis on
longitudinal data, our method can be used to further inform the
iterative algorithm of the available template information. The scheme
is presented in Fig.~\ref{fig:Lit}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{../images/Lit.png}
  \caption{Illustration of the application of our method (marked
    in yellow, red) within the deep learning based framework of
    \cite{kim2019extreme} (in blue).}
  \label{fig:Lit}
\end{figure}

 %Hence, one may in principle use this full sinogram as an alternate initial reconstruction of the test within our framework.

Our focus in this paper instead is an analytical method that uses the
available longitudinal data to better inform the reconstruction of the
test.  We present a technique that makes use of multiple object
priors, corresponding to high quality reconstructions of similar
objects across time, typically in a longitudinal study.  We use these
priors to improve the reconstruction of the test from largely
undersampled measurements. In particular, we define regions of change
between the test and the eigenspace spanned by previous high quality
object-priors in a purely data driven manner.  We first construct a
`weights map' that gives clear insights about regions of change within
the object over time.  Additionally, our technique differentiates
between genuine structural changes and ``changes'' that appear due to
undersampling artefacts.  Besides this, since we use a statistical
model with multiple object-priors, we avoid the problem of selecting
an appropriate single object-prior
unlike~\cite{PICCS,Pourmorteza2015,Lee2012,pirple,Marjolein2016} and
also make use of the additional information available in multiple
priors. Our technique also has the advantage of not requiring any
expert-domain knowledge about the object being scanned.

@@

In this section, we first illustrate the advantage of using multiple
types of eigenspaces for the computation of weights map, and then
present reconstruction results on all datasets.  We show both
qualitative and quantitative results.  Before we do this, it is important
to first consider our method of reporting our results.
